---
subtitle: "TMA4268 Statistical Learning V2022"
title: "Compulsory exercise 1: Group 16"
author: "Weicheng Hua, Emil Johannesen Haugstvedt, Torbjørn Baadsvik"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE,tidy=TRUE,message=FALSE,warning=FALSE,strip.white=TRUE,prompt=FALSE,
                      cache=TRUE, size="scriptsize",fig.width=4, fig.height=3,fig.align = "center")

```

```{r,eval=TRUE,echo=FALSE, warning=FALSE}
library("knitr")
library("rmarkdown")

# Classification
library(e1071)
library(caret)
library(kernlab)

# Partial least squares regression
library(plsRglm)
library(plsdof)

# K-means clustering
#install.packages(c("factoextra", "cluster"))
library(factoextra)
library(cluster)
```


\section{Problem 1}

\subsection{a)}

\subsection{b)}

\subsection{c)}

\subsection{d)}


\section{Problem 2}

\subsection{a)}

\subsection{b)}


\section{Problem 3}

\subsection{a)}

\subsection{b)}


\section{Problem 4}

\subsection{a)}

\subsection{b)}

\subsection{c)}
```{r,eval=TRUE,echo=TRUE, warning=FALSE}
library(tidyverse)
library(palmerpenguins) # Contains the data set "penguins".
data(penguins)
names(penguins) <- c("species","island","billL","billD","flipperL","mass","sex","year")
Penguins_reduced <- penguins %>%  
  dplyr::mutate(mass = as.numeric(mass),  
         flipperL = as.numeric(flipperL),
         year = as.numeric(year)) %>% 
  drop_na()
# We do not want "year" in the data (this will not help for future predictions)
Penguins_reduced <- Penguins_reduced[,-c(8)]
set.seed(4268)
# 70% of the sample size for training set
training_set_size <- floor(0.7 * nrow(Penguins_reduced))
train_ind <- sample(seq_len(nrow(Penguins_reduced)), size = training_set_size)
train <- Penguins_reduced[train_ind, ]
test <- Penguins_reduced[-train_ind, ]
```

\subsection{d)}

\section{Problem 5}

\subsection{a)}
\begin{tabular}{c c c c}
i) & ii) & iii) & iv)\\
FALSE & TRUE & FALSE & TRUE\\
\end{tabular}

\subsection{b)}

```{r}

svc.cvtune <- function(kernel, paramgrid, k){
 ctrl <- tune.control(sampling="cross", cross=k, nrepeat = 1)
 cvtune.result <- tune(method=svm, species~., kernel=eval(kernel), 
      data=train, ranges=paramgrid, tunecontrol=ctrl)
 cvtune.result
}

k <- 5
svc.fit_and_eval <- function(kernel, paramgrid, k){
  res <- svc.cvtune(kernel, paramgrid, k)
  print(strrep("_", 60))
  print(paste(eval(kernel), "support vector classifier"))
  print(strrep("-", 60))
  print("Parameters:")
  print(c(res$best.parameters))
  
  model <- res$best.model
  pred <- as.factor(predict(model, test[-c(1)]))
  cm <- confusionMatrix(data=pred, reference=test$species)
  print(cm)
}

svc.fit_and_eval("linear", data.frame(cost=.1*1:20), k)
svc.fit_and_eval("radial", data.frame(cost=.1*1:20, gamma=.1*1:20), k)
```
Using a linear rather than radial kernel in the support vector classifier
yields slightly superior results on the test set. Thus, the linear kernel is
preferred.


\section{Problem 6}

\subsection{a)}
```{r}
id <- "1NJ1SuUBebl5P8rMSIwm_n3S8a7K43yP4" # google file ID
happiness <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", id),fileEncoding="UTF-8-BOM")
colnames(happiness)
```

```{r}

cols = c('Country.name', 
         'Ladder.score',  # happiness score
         'Logged.GDP.per.capita',  
         'Social.support', 
         'Healthy.life.expectancy', 
         'Freedom.to.make.life.choices',
         'Generosity',  # how generous people are
         'Perceptions.of.corruption')
# We continue with a subset of 8 columns:
happiness = subset(happiness, select = cols)
rownames(happiness) <- happiness[, c(1)]
# And we create an X and a Y matrix
happiness.X = happiness[, -c(1, 2)]
happiness.Y = happiness[, c(1, 2)]
happiness.XY = happiness[, -c(1)]
# scale
happiness.X = data.frame(scale(happiness.X))
str(happiness)
```


```{r, fig.height=9, fig.width=10,out.width='18cm'}
library(ggfortify)
pca_mat = prcomp(happiness.X, center=T, scale=T)
# Score and loadings plot:
autoplot(pca_mat, data = happiness.X, colour='Black',
         loadings = TRUE, loadings.colour = 'red',
         loadings.label = TRUE, loadings.label.size = 5, 
         label=T, label.size=4.5)
```
\subsubsection{i)}
We observe that the variables "Logged.GDP.per.capita", "Healthy.life.expenctancy"
and "Social.support" are highly correlated as they have nearly equal loadings on
PC1 and PC2.
The loading vectors for the variables "Freedom.to.make.like.choices" and 
"Perceptions.of.corruption" are nearly antiparallel, indicating that these
variables have a strong negative correlation.

\subsubsection{ii)}
Afghanistan appears to be clearly separated from the other countries in this
plot, and may be considered to be an outlier.


\subsection{b)}

\subsubsection{i)}
```{r}
rot <- pca_mat$rotation
loading.PC1 <- data.frame(variable=rownames(rot), loading=abs(rot[,1]))
ggplot(data=loading.PC1, aes(x=variable, y=loading)) +
  geom_col(color="white", fill="red") +
  coord_flip()
```

\subsubsection{ii)}
```{r}
summary(happiness.XY)
plsr_model.1 <- plsR(formula=Ladder.score~., data=happiness.XY, 
                   scaleX=T, scaleY=T, nt=1)
plsr_model.1

```

\subsubsection{iii)}
```{r}
rot.PC1 <- plsr_model.1$ww
rot.PC1
rownames(rot.PC1)
loading.PC1 <- data.frame(variable=rownames(rot.PC1), loading=abs(rot.PC1[,1]))
loading.PC1
ggplot(data=loading.PC1, aes(x=variable, y=loading)) +
  geom_col(color="white", fill="red") +
  coord_flip()
```

\subsubsection{iv)}
Based on the PLSR we see that the variables "Logged.GDP.per.capita", 
"Healthy.life.expectancy", and "Social.support" are the most important 
predictors.


\subsection{c)}

\begin{tabular}{c c c c}
i) & ii) & iii) & iv)\\
FALSE & FALSE & TRUE & TRUE\\
\end{tabular}

\subsection{d)}
\subsubsection{i)}
```{r}
K = 3
km.out <- kmeans(happiness.X, K, iter.max = 10)
invalidclustering <- function(km.out){
  clust <- km.out$cluster
  scand <- clust[c("Norway", "Denmark", "Sweden", "Finland")]
  !((sd(scand) == 0) & (clust["United States"] != scand["Norway"]))
}

i <- 1
print(paste("iteration", i))
while(invalidclustering(km.out)){
  i <- i + 1
  print(paste("iteration", i))
  km.out <- kmeans(happiness.X, K, iter.max = 10)
}

autoplot(pca_mat, data = happiness.X, colour=km.out$cluster,
         label=F, label.size=5,
         loadings = F, loadings.colour = 'blue',
         loadings.label = F, loadings.label.size = 3)
```
K = 3 was the minimum parameter required to obtain the desired clustering conditions.

\subsubsection{ii)}
```{r}
clust <- km.out$cluster
clust[c("Norway", "Denmark", "Sweden", "Finland")]
clust["United States"]
mean_ladder <- sapply(1:3, function(i) mean(happiness.Y[names(clust[clust == i]),2]))
mean_ladder
```
We observe that the clustering algorithm places the scandinavian countries in 
cluster 1, and the United States in cluster 2.
When computing the average value of the "Ladder.score" variable within each cluster,
we find that "Ladder.score" decreases with increasing cluster index.
Thus, we conclude that the scandinavian countries belong the "happiest"
cluster, and that the United States belong to the "medium happiness" cluster.






