---
subtitle: "TMA4268 Statistical Learning V2022"
title: "Compulsory exercise 1: Group 16"
author: "Weicheng Hua, Emil Johannesen Haugstvedt, Torbjørn Baadsvik"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE,tidy=TRUE,message=FALSE,warning=FALSE,strip.white=TRUE,prompt=FALSE,
                      cache=TRUE, size="scriptsize",fig.width=4, fig.height=3,fig.align = "center")

```

```{r,eval=TRUE,echo=FALSE}
library("knitr")
library("rmarkdown")
```

\section{Problem 1}

\subsection{a)}

\subsection{b)}

\subsection{c)}

\subsection{d)}

\section{Problem 2}

\subsection{a)}

\subsection{b)}

\section{Problem 3}

\subsection{a}

(i) True. The proof for this is pretty long and complicated. You can see it in the link

(ii) False. By increasing the number of cutpoints each step function will get more and more affected by the points within their range, and thus overfit.

(iii) False. The penalty term is \$\\int g''(t)\^2 dt\$.

(iv) True. With high \$k\$ the number more neighbors are needed in order to classify a point, thus the variance will be low and the bias will increase.

\subsection{b}

``` {r}
# Fit model
model <- gam(medv ~ rm + s(ptratio, k = 3) + poly(lstat, df = 2), data = boston.train)

# Plot model with training data
plot(model, boston.train)
```

\section{Problem 4}

\subsection{a)}

(i) False. The trees can handle interaction terms, but you can not specify them yourself. The tree will "find" them by its nature.

(ii) True. 

(iii) True.

(iv) True. It decides the number of "iterations" in the boosting algorithm.

\subsection{b)}


\subsection{c)}
First some given R code
```{r}
library(tidyverse)
library(palmerpenguins) # Contains the data set "penguins".
data(penguins)

names(penguins) <- c("species","island","billL","billD","flipperL","mass","sex","year")

Penguins_reduced <- penguins %>%  
  dplyr::mutate(mass = as.numeric(mass),  
         flipperL = as.numeric(flipperL),
         year = as.numeric(year)) %>% 
  drop_na()

# We do not want "year" in the data (this will not help for future predictions)
Penguins_reduced <- Penguins_reduced[,-c(8)]

set.seed(4268)
# 70% of the sample size for training set
training_set_size <- floor(0.7 * nrow(Penguins_reduced))
train_ind <- sample(seq_len(nrow(Penguins_reduced)), size = training_set_size)
train <- Penguins_reduced[train_ind, ]
test <- Penguins_reduced[-train_ind, ]
```

(i)
```{r}
library(tree)

# Create the tree 
tree.penguin <- tree(species ~ .,
                     data = train,
                     split = 'gini')

# Plot the tree
plot(tree.penguin, type = 'uniform');text(tree.penguin)
```
(ii)
Now, apply 10-fold cross-validation 
```{r}
set.seed(123)

# Do 10-fold cross-validation
cv.penguin <- cv.tree(tree.penguin)

# Plot result from the 10-fold cross-validation
plot(cv.penguin)
```
(iii) 
From the above plot the optimal tree looks to be of size 4. 
```{r}

# Prune the tree according to the obseravtion in the above plot
prune_tree.penguin <- prune.misclass(tree.penguin, best = 4)

plot(prune_tree.penguin, type = 'uniform'); text(prune_tree.penguin)
```


```{r}
pred <- predict(prune_tree.penguin, test, type = 'class')
tab <- table(pred, test$species)
tab
```
The misclassification rate will then be:
```{r}
(1 - sum(diag(tab))/sum(tab))
```


\subsection{d)}
```{r}
library(gbm)
library(dismo)

#train.df <- data.frame(train)
#train.df['isAdelie'] <- train.df$species == 'Adelie'
#train.df['isChinstrap'] <- train.df$species == 'Chinstrap'
#train.df['isGentoo'] <- train.df$species == 'Gentoo'


#
#a <- gbm.step(data = train.df,
#              gbm.x = names(train.df[2:7]),
#              gbm.y = names(train.df[8:10]),
#              family = 'gaussian',
#              tolerance.method = 'auto')

boosted.penguin <- gbm(species~ .,
                       data = train,
                       n.trees = 500,
                       distribution = 'multinomial')

summary(boosted.penguin)
```

Now we predict on the test data and calculate the misclassification rate:

```{r}
pred <- predict.gbm(boosted.penguin, test, type = 'response')

pred.penguin <- colnames(pred)[apply(pred,1,which.max)]

tab.boost <- table(pred.penguin, test$species)
tab.boost
```
The misclassification rate is
```{r}
misclassification.rate.boost <- 1-sum(diag(tab.boost))/sum(tab.boost)
misclassification.rate.boost
```


\section{Problem 5}

\subsection{a)}

\subsection{b)}

\section{Problem 6}

\subsection{a)}

\subsection{b)}

\subsection{c)}

\subsection{d)}
